---
title: "Homework 3"
author: "Joshua Carpenter"
date: "2023-10-14"
output: github_document
---

```{r setup, results='hide', message=FALSE, warning=FALSE}
library(tidyverse)
library(p8105.datasets)
library(scales)
library(knitr)
library(janitor)
data("instacart")
data("brfss_smart2010")
```

## Problem 1
The `instacart` dataset contains order history from the online shopping service Instacart in 2017. The dataset has one row per item per order and contains info about when the order was placed, whether the customer had bought the item before, and which aisle the item came from. Here are a few rows from the dataset:

```{r echo=FALSE}
instacart %>%
  filter(aisle_id <= 6) %>%
  slice_head(n = 1, by = aisle) %>%
  kable()
```

I just realized this problem is worth zero points, so I'm not going to put any more work into it, but here are some nice tables and figures:
```{r fig.asp=1.8, message=FALSE}
# Bar chart
instacart %>%
  group_by(aisle_id, aisle) %>%
  count() %>%
  arrange(n) %>%
  filter(n >= 10000) %>%
  ggplot(mapping = aes(x = n, y = factor(aisle, levels = aisle))) +
  geom_col(width = 0.7, fill = '#ea9905') +
  geom_text(aes(label = comma(n)), hjust = -0.3) +
  theme_void() +
  theme(axis.text.y = element_text(hjust = 0, 
                                   margin = margin(r = 10, l = 10))) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.2))) +
  xlab("") +
  ylab("") +
  ggtitle("Number of Items Ordered", "From Aisles with at Least 10,000 Orders")
```

#### Top Ordered Items from Select Aisles
```{r message=FALSE}
# Table
instacart %>%
  group_by(aisle_id, aisle, product_id, product_name) %>%
  filter(aisle %in% c("baking ingredients", "dog food care", 
                      "packaged vegetables fruits")) %>%
  count() %>%
  ungroup() %>%
  slice_max(n, by = aisle_id, n = 3) %>%
  select(-ends_with("_id")) %>%
  kable(col.names = c("Aisle", "Product", "Number Ordered"))
```

#### Average Time of Day (hours since midnight) for Orders of Coffee Ice Cream and Pink Lady Apples
```{r message=FALSE}
# Assuming that 0 is Sunday, since the data description doesn't specify
dow <- c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", 
         "Friday", "Saturday")

instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarise(mean = round(mean(order_hour_of_day), 1)) %>%
  mutate(order_dow = dow[order_dow + 1]) %>%
  pivot_wider(names_from = product_name, values_from = mean) %>%
  rename(`Day of Week` = order_dow) %>%
  kable()
```

## Problem 2
```{r message=FALSE}
brfss_clean <- brfss_smart2010 %>%
  clean_names() %>%
  mutate(response = factor(response, 
                           levels = c("Poor", "Fair", "Good", 
                                      "Very good", "Excellent"),
                           ordered = TRUE)) %>%
  filter(topic == "Overall Health", !is.na(response)) %>%
  rename(state = locationabbr, location = locationdesc)
```

In 2010, there were quite a few more states with more than 7 locations than in 2002. The only state that had more than 7 locations in 2002 but didn't by 2010 was Connecticut.
```{r}
brfss_clean %>%
  group_by(year, state) %>%
  summarise(n = n_distinct(location)) %>%
  filter(year %in% c(2002, 2010), n >= 7) %>%
  pivot_wider(names_from = year, values_from = n, id_cols = state) %>%
  kable()
```

This spaghetti plot shows us that there aren't a lot of clear patterns for average value over time and that there aren't many outlying state averages over this period of time.
```{r}
brfss_state_avg <- brfss_clean %>%
  filter(response == "Excellent") %>%
  group_by(year, state) %>%
  summarise(avg_val = mean(data_value, na.rm = TRUE))
  
ggplot(brfss_state_avg, mapping = aes(x = year, y = avg_val, group = state)) +
  geom_line() +
  theme_bw() +
  scale_x_continuous(expand = expansion(0)) +
  xlab("Year") +
  ylab("Average Value by State")
```

Between 2006 and 2010 there seems to be a slight rightward shift in the distribution. In 2010, quite a few more people answered "Very good" compared to 2006.
```{r}
brfss_clean %>%
  filter(year %in% c(2006, 2010), state == "NY") %>%
  group_by(year, response) %>%
  summarise(mean = mean(data_value)) %>%
  ggplot(aes(x = response, y =  mean)) +
  geom_col() +
  facet_grid(~year) +
  theme_bw() +
  xlab("Response") +
  ylab("Average Value for NY")
```

## Problem 3
```{r}
dm <- 
  read_csv("Data/nhanes_covar.csv", skip = 4) %>%
  clean_names() %>%
  mutate(sex = factor(sex, labels = c("Male", "Female")),
         education = factor(education, ordered = TRUE,
                            labels = c("Less than high school",
                                                  "High school equivalent",
                                                  "More than high school"))) %>%
  filter(age >= 21)
accel <- 
  read_csv("Data/nhanes_accel.csv") %>%
  pivot_longer(starts_with("min"), names_to = "min", values_to = "accel") %>%
  rename(seqn = SEQN) %>%
  mutate(min = parse_integer(substring(min, 4))) %>%
  right_join(dm, by = "seqn") %>%
  select(seqn, sex, age, bmi, education, min, accel)
```

Age seems to have a fairly wide spread. We have a higher proportion of people who have completed more than high school compared to the other education categories. Demographics are similar between genders.
```{r}
dm %>%
  group_by(sex, education) %>%
  count() %>%
  mutate(education = replace_na(as.character(education), "Unknown")) %>%
  pivot_wider(names_from = sex, values_from = n) %>%
  rename(Education = education) %>%
  kable()

ggplot(dm, mapping = aes(x = age)) +
  geom_histogram(bins = 15) +
  facet_grid(~sex) +
  xlab("Age") +
  ylab("Count")
```

More education is associated with staying more physically active later in life. In general, females are more active than males, except for less educated females over 50 years old.
```{r fig.asp=2}
accel %>%
  group_by(seqn, sex, age, bmi, education) %>%
  summarise(accel = sum(accel)) %>%
  ggplot(mapping = aes(x = age, y = accel, color = sex)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_grid(rows = vars(education)) +
  labs(x = "Age", y = "Movement", color = "Sex")
```

There are a lot of high outliers for the highest education group, including some in the evening that are cut off in order to see the trend better. That pattern seems indicative of "working out" or intentional exercise. The lowest education group seems to have heaveir physical activity in the morning - potentially indicative of a physically active job.
```{r fig.asp=2}
accel %>%
  ggplot(mapping = aes(x = min, y = accel, color = sex)) +
  geom_point(alpha = 0.3, size = 0.5) +
  geom_smooth(se = FALSE) +
  ylim(c(0, 50)) +
  facet_grid(rows = vars(education)) +
  labs(x = "Minute", y = "Movement", color = "Sex") +
  theme_bw()
```

